---
title: "IPEDS World Domination Guide"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{IPEDS World Domination Guide}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)
# need to add all packages that we are using here
pkgs <- c("tidyverse", "readxl", "RCurl", "data.table", "stringr")
for(pkg in pkgs) {
  if(!require(pkg, character.only = TRUE)) {
    install.packages(pkg)
    library(pkg, character.only = TRUE)
  }
  library(pkg, character.only = TRUE)
}


path <- ifelse(file.exists("S:/"), "S:/", "/Volumes/files/Shared/")
# setwd(path)


#this one merges files together
source("https://raw.githubusercontent.com/emmamorgan-tufts/IPEDS_longitudinal/master/merge_ipeds_data.R")

#this one adds value labels when appropriate
source("https://raw.githubusercontent.com/emmamorgan-tufts/IPEDS_longitudinal/master/add_valuesets.R")

#this one changes varnames to english titles (after values have been addressed)
source("https://raw.githubusercontent.com/emmamorgan-tufts/IPEDS_longitudinal/master/change_varnames_to_vartitles.R")
```

# Goal/Summary

Run all ipeds surveys through our IWD system to data check whether they will compile.


# Data Sources

1) IPEDS surveys. The root directory is S:/IRO/resources/IPEDS/All Surveys

2) Peer list


# Set up
## Packages
```{r}
pkgs <- c("tidyverse", "readxl", "RCurl", "data.table", "stringr")
for(pkg in pkgs) {
  if(!require(pkg, character.only = TRUE)) {
    install.packages(pkg)
    library(pkg, character.only = TRUE)
  }
  library(pkg, character.only = TRUE)
}
``` 

Functions for peerlist clean up

```{r}
script_peerList <- RCurl::getURL("https://raw.githubusercontent.com/emmamorgan-tufts/IPEDS_longitudinal/master/peerList.R", 
                                 ssl.verifypeer = FALSE)
eval(parse(text = script_peerList))
rm("script_peerList")
```

Functions for compiling
```{r}
script_compile_IPEDS <- RCurl::getURL("https://raw.githubusercontent.com/emmamorgan-tufts/IPEDS_longitudinal/master/compile_IPEDS_survey.R", 
                                      ssl.verifypeer = FALSE)
eval(parse(text = script_compile_IPEDS))
rm(script_compile_IPEDS)
```

## File Structure
Data and Dictionary


# Set Up a Peerlist

```{r peerlist}
peer_filepath <- "myfile.csv"
IPEDS_peers <- IPEDS_peers_from_file(peer_filepath)
```

# Create file location
```{r location}
IPEDS_data_location_general <- "IRO/resources/IPEDS/All Surveys"

surveyFolder <- "Admissions"
surveySection <- "adm"

IPEDS_data_location <- paste(path, IPEDS_data_location_general,surveyFolder,surveySection, sep="/")

script_compile_IPEDS <- RCurl::getURL("https://raw.githubusercontent.com/emmamorgan-tufts/IPEDS_longitudinal/master/compile_IPEDS_survey.R", 
                                      ssl.verifypeer = FALSE)
eval(parse(text = script_compile_IPEDS))
rm(script_compile_IPEDS)

```

# Merge files
```{r merge}  
IPEDS_merge <- merge_IPEDS_data(IPEDS_data_location, peer_UNITIDs = peerlist)
IPEDS_data <- IPEDS_merge$data
IPEDS_dictionary <- IPEDS_merge$dictionary
IPEDS_valuesets <- IPEDS_merge$valuesets
glimpse(IPEDS_dictionary)
glimpse(IPEDS_valuesets)
```

# Add values
```{r values}
IPEDS_data_values <- add_values(longtable = IPEDS_data, valueset = IPEDS_valuesets, ignore_size_warning = T)
```

# Add variable labels
```{r variables}
IPEDS_data_clean <- change_varnames_vartitles(longtable = IPEDS_data_values, varnames = IPEDS_dictionary, ignore_size_warning = T)
glimpse(IPEDS_data_clean)
```


# Output files
```{r writeout}
# REMEBER TO CREATE A COMPILED FILE IN SURVEY FOLDER.
write.csv(IPEDS_data_clean,  paste0(IPEDS_data_location,"/compiled/",surveySection, "_compiled.csv"), row.names = F)

write.csv(IPEDS_dictionary,  paste0(IPEDS_data_location,"/compiled/",surveySection, "_dictionary.csv"), row.names = F)

write.csv(IPEDS_valuesets,  paste0(IPEDS_data_location,"/compiled/", surveySection, "_valuesets.csv"), row.names = F)

```

