---
title: "IPEDS World Dominiation Data Compilation"
author: "KMF"
date: "updated March 2024 to be paramterized"
params: 
  survey: 12 Month Enrollment
  subsurvey: effy
---


```{r setup}

knitr::opts_chunk$set(include = FALSE)

pkgs <- c("tidyverse", "readxl", "RCurl", "knitr")
for(pkg in pkgs) {
  if(!require(pkg, character.only = TRUE)) {
    install.packages(pkg, repos = "http://cran.us.r-project.org")
    library(pkg, character.only = TRUE)
  }
  library(pkg, character.only = TRUE)
}


path <- ifelse(file.exists("S:/"), "S:/", "/Volumes/files/Shared/")
# setwd(path)


#this one merges files together
source("https://raw.githubusercontent.com/emma-morgan/IPEDS_longitudinal/master/merge_ipeds_data.R")


#this one adds value labels when appropriate
source("https://raw.githubusercontent.com/emma-morgan/IPEDS_longitudinal/master/add_valuesets.R")


#this one changes varnames to english titles (after values have been addressed)
source("https://raw.githubusercontent.com/emma-morgan/IPEDS_longitudinal/master/change_varnames_to_vartitles.R")

```

# Goal/Summary

Run all ipeds surveys through our IWD system to data check whether they will compile.


# Data Sources

1) IPEDS surveys. The root directory is now S:/IRO/resources/IPEDS/IWD - Data

2) Peer list -- in this case peer list is null because we are trying FULL compile

# Read in Peerlist

```{r peerlist}
#peerlist <- read.csv(paste0(path, "/IRO/resources/IPEDS/Simple_Peer_List.csv"))

peerlist <- NULL
```

# Create file location
```{r location, include=FALSE}
  
IPEDS_data_location_general <- paste0("S:/IRO/resources/IPEDS/IWD - Data")

surveyFolder <- params$survey
surveySection <- params$subsurvey

IPEDS_data_location <- paste(IPEDS_data_location_general,surveyFolder,surveySection, sep="/")
```

# Merge files
```{r merge}  
IPEDS_merge <- merge_IPEDS_data(IPEDS_data_location, peer_UNITIDs = peerlist$UNITID)
IPEDS_data <- IPEDS_merge$data
IPEDS_dictionary <- IPEDS_merge$dictionary
IPEDS_valuesets <- IPEDS_merge$valuesets
glimpse(IPEDS_dictionary)
glimpse(IPEDS_valuesets)

```


in order to do full compile, need to use loop to break file apart, call the values and varnames functions, and then bind file back together. (otherwise computer meory is insufficient to run the two functions)
```{r}

IDS <- unique(IPEDS_data$UNITID)
x = length(IDS)

dslist <-  list()

for (i in seq(1, x, 1000)) {
  
  ds <- IPEDS_data %>% filter(UNITID %in% IDS[i:(i+999)])
  IPEDS_data_values <-  add_values(longtable = ds, valueset = IPEDS_valuesets, ignore_size_warning = T)
  IPEDS_data_clean <- change_varnames_vartitles(longtable = IPEDS_data_values, varnames = IPEDS_dictionary, ignore_size_warning = T)
  
  dslist[[i]] <- assign(paste0("ds", i), IPEDS_data_clean) 
  print(i)
  
}

full <- bind_rows(dslist)


```



# Output files
```{r writeout}
# REMEBER TO CREATE A COMPILED FILE IN SURVEY FOLDER.
write_csv(full,  paste0(IPEDS_data_location,"/compiled/",surveySection, "_compiled_full.csv"))

write.csv(IPEDS_dictionary,  paste0(IPEDS_data_location,"/compiled/",surveySection, "_dictionary_full.csv"), row.names = F)

write.csv(IPEDS_valuesets,  paste0(IPEDS_data_location,"/compiled/", surveySection, "_valuesets_full.csv"), row.names = F)

```

```{r}
rm(list = ls()[grep("^ds", ls())])
```

